Lawvere-Tierney Sheafification\\ in Homotopy Type Theory

### 2015-10-12 paper/sheaf_lics.tex
finish Simon's proof

The Giraud-Rezk-Lurie axioms are the $\infty$-version of Giraud's
axioms that characterize a topos. Namely, there are 4 axioms on a
$(\infty,1)$-category that have been shown to be equivalent to
$(\infty,1)$-topos~\cite[Chapter 6]{lurie}. This is not the purpose of
this paper to discuss those axioms, we only state here a particular
case of the last axiom which says that every groupoid object is
effective.
%
This particular case connects a surjection (or a $(-1)$-connected
function) to the colimit of its \v{C}ech nerve.
%
This is the $\infty$-generalization of the fact that, in a topos,
every epimorphism is the colimit of its kernel pair, which plays a
central part in the definition of sheafification.

\begin{defi}
  Let $f:X \to Y$ be a map, and $p:\nat$. The $p$-pullback of $f$
  (also called the $p$-fold fiber product of $X$ over $Y$),
  noted $X\times_Y \cdots\times_Y X$, is
  the limit of the diagram
  $$\xymatrix{
    X \ar[rd]^f & \cdots & X \ar[ld]_f \\
      &    Y   &
  }$$
  with $p$ copies of $X$. The $0$-pullback of $f$ is $\one$.
\end{defi}

As explained in Section~\ref{sec:colim-homot-type}, every limit exists
in (homotopy) type theory and can be described by a $\Sigma$-type. 
%
The particular case of a $p$-pullback is given by
%
$$
X\times_Y \cdots\times_Y X = \sum_{x:X^n} (f x_1 = f x_2) \land
\cdots \land (f x_{n-1} = f x_n).
$$
%
The \v{C}ech nerve of a map $f$ can then be described as the
simplicial object that, at degree $p$, is given by the
$(p+1)$-pullback of $f$. The \v{C}ech nerve can thus also be described
as a diagram over $\Delta^{op}$ (where $\Delta$ is the simplicial
category, seen as a graph). For simplicity, we only give an intuitive
definition here, the complete definition can be found in our
formalization.
%
\begin{defi}
  Let $f$ be a map from $X$ to $Y$. The {\em \v{C}ech nerve} $C(f)$ of $f$
  is given by the diagram
  $$\xymatrix{
    C(f) := \cdots~ X \times_Y X \times_Y X \tar[r]{4pt} & X \times_Y X \dar[r]{2pt} & X
  }$$
where the arrows are the canonical projections.
\end{defi}

\begin{ax}[Giraud-Rezk-Lurie]
  For any surjection $f : X \to Y$, the colimit of its \v{C}ech nerve
  $C(f)$ is $Y$.
\end{ax}

With the point of view that homotopy type theory is a type-theoretic
version of higher topos theory, Theorem 6.1.0.6 of~\cite{lurie}
suggests that this axiom can be proved in homotopy type theory. 
%
But proving this axiom seems to be a quite complex task on its own,
and we have decided to admit it in this paper.

@@

\begin{lem}\label{lem:sepiscolim}
  Let $T:\Type_{n+1}$. Then $\separated T$ is the colimit of the closed
  diagonal diagram
  $$\xymatrix @C=4em  { 
    \cdots \overline{\Delta_3} \tar[r]{4pt} & \overline{\Delta_2} \dar[r]{2pt} &
    \overline{\Delta_1}
  }$$
where $\Delta_k$ is the $k$-pullback of $\id : T \to T$.
\end{lem}

@@

\begin{proof}
  As $\mu_T$ is an surjection, using the particular case of 
  Giraud-Rezk-Lurie axiom introduced in Section~\ref{sec:giraud-ax},
  we know that $\separated T$ is the colimit of $C(\mu_T)$. 
%
  Thus the result follows from the fact that $C(\mu_T) = C(\id)$,
  which we now show.

  Let $k:\nat$, we have to show that 
  \begin{align*}
    &\sum_{t:T^k} (\mu_T t_1 = \mu_T t_2 \land \cdots
      \land \mu_T t_{k-1} \mu_T t_k)\\
    &= \sum_{t:T^k} \modal_n (t_1 = t_2 \land \cdots
      \land t_{k-1} = t_k)
  \end{align*}
By induction on $k$, and the preservation of products by $\modal_n$, it
suffices to show that for any $a,b:T$, $\modal_n (a=b) = (\mu_T a =
\mu_T b)$. By univalence, we want arrows in both ways, forming an
equivalence.
\begin{itemize}
\item Suppose $p : (\mu_T a = \mu_T b)$. Then projecting $p$ along
  first components yields $q : \prod_{t:T} \modal_n(a=t) = \modal_n (b=t)
  $.
  Taking for example $t=b$, we deduce $\modal_n (a=b) = \modal_n(b=b)$,
  and the latter is inhabited by $\eta_{b=b} 1$.
\item Suppose now $p : \modal_n(a=b)$. Let $\iota$ be the first
  projection from $\separated T \to (T \to \Type_n^\modal)$. $\iota$ is
  an embedding, thus it suffices to prove $\iota (\mu_T a) = \iota
  (\mu_T b)$, \ie $\prod_{t:T}\modal_n (a=t) = \modal_n (b=t)$. The latter
  remains true by univalence.
\end{itemize}
The fact that these two form an equivalence is technical, we refer to
the formalization for an explicit proof.
\end{proof}

>>>>

The Giraud-Rezk-Lurie axioms are the $\infty$-version of Giraud's
axioms that characterize a topos. Namely, there are 4 axioms on a
$(\infty,1)$-category that have been shown to be equivalent to
$(\infty,1)$-topos~\cite[Chapter 6]{lurie}. In~\cite{boulier}, a
type-theoritic version of these axioms are proposed.
%
This particular case connects a surjection (or a $(-1)$-connected
function) to the colimit of its \v{C}ech nerve.
%
This is the $\infty$-generalization of the fact that, in a topos,
every epimorphism is the colimit of its kernel pair, which plays a
central part in the definition of sheafification.

\begin{defi}
  Let $f:X \to Y$ be a map. $T_f$ is the higher inductive type given
  by
  \begin{itemize}
  \item $t:~X \to T_f$
  \item $\alpha:~\forall a\,b:X,~f a = f b \to t a = t b$
  \item $\alpha_1:~\forall a:X,~\alpha\, a\, a\, 1 = 1$
  \end{itemize}
  We view $T_f$ as the coequalizer of
  $$\xymatrix{
    \sum_{a,b:X} f a = f b \dar[r]{4pt}^-{\pi_1}_-{\pi_2} & X
  }$$
  preserving the identity.

  We call $\tilde f$ the map $T_f \to Y$ given by induction.
\end{defi}
%
The \v{C}ech nerve of a map $f$ can then be described as the
simplicial object that, at degree $p$, is given by the
$p$th iteration of $T$. 
%
\begin{defi}
  Let $f$ be a map from $X$ to $Y$. The {\em \v{C}ech nerve} $C(f)$ of $f$
  is given by the diagram
  $$\xymatrix{
    C(f) := X \ar[r]^t & T_f \ar[r]^t & T_{\tilde f} \ar[r] &\cdots
% \cdots~ X \times_Y X \times_Y X \tar[r]{4pt} & X \times_Y X \dar[r]{2pt} & X
  }$$
\end{defi}

\begin{thm}[Giraud-Rezk-Lurie]
  For any surjection $f : X \to Y$, the colimit of its \v{C}ech nerve
  $C(f)$ is $Y$.
\end{thm}

With the point of view that homotopy type theory is a type-theoretic
version of higher topos theory, this seems to be the analogous of Theorem 6.1.0.6 of~\cite{lurie}.
%
The proof of this theorem can be found in~\cite{boulier}.

@@

\begin{defi}
  Let $X:\Type$. Let $\mathring T_X$ be the higher inductive type
  generated by
  \begin{itemize}
  \item $\mathring t:~\|X\|_{n+1} \to \mathring T_X$
  \item $\mathring \alpha:~\forall a\, b:\|X\|_{n+1},~\modal (a=b) \to
    \mathring t a = \mathring t b$
  \item $\mathring \alpha_1:~\forall a:\|X\|_{n+1},~
    \mathring \alpha\, a \, a\, (\eta_{a=a} 1) = 1$
  \end{itemize}

  We view $\mathring T$ as the coequalizer of
  $$\xymatrix{\sum_{a,b:\|X\|_{n+1}} \modal (a=b) \dar[r]{4pt}^-{\pi_1}_-{\pi_2}
  & \|X\|_{n+1}}$$
  preserving $\eta_{a=a} 1$.

  We consider the diagram $\mathring T$ :
  $$\xymatrix{\|X\|_{n+1} \ar[r] & \|\mathring T_{X}\|_{n+1} \ar[r] & \|\mathring
  T_{\mathring T_X} \|_{n+1} \ar[r] & \cdots} $$
\end{defi}

\begin{lem}\label{lem:sepiscolim}
  Let $T:\Type_{n+1}$. Then $\separated T$ is the $(n+1)$-colimit of the
  diagram $\mathring T$.
\end{lem}

@@

We will need the following lemma:

\begin{lem}
  Let $A,S:\Type_{n+1}$, $S$ separated, and $f:A \to S$. Then if 
  $$\forall a,b:A,~f a = f b \simeq \modal (a=b),$$ then 
  $$\forall a,b:\|T_f\|_{n+1},~|\tilde f|_{n+1} a = |\tilde f|_{n+1} b \simeq \modal (a=b).$$
\end{lem}

\begin{proof}[Sketch of proof]
  By induction on truncation, we need to show that 
  $$\forall a,b:T_f,~\tilde f |a|_{n+1} = \tilde f |b|_{n+1} \simeq
  \modal (|a|_{n+1}=|b|_{n+1}).$$

  We use the encode-decode method to characterize $\tilde f |a|_{n+1}
  = x$, and the result follows. We refer to the formalization for details.
\end{proof}

\begin{proof}[Proof of lemma~\ref{lem:sepiscolim}]
  As $\mu_T$ is an surjection, using the particular case of 
  Giraud-Rezk-Lurie axiom introduced in Section~\ref{sec:giraud-ax},
  we know that $\separated T$ is the $(n+1)$-colimit of $\|C(\mu_T)\|_{n+1}$. 
%
  Thus the result follows from the fact that $\|C(\mu_T)\|_{n+1} =
  \mathring T$,
  which we now show.

  By induction, we need to show that for any $A,B:\Type_{n+1}$ such
  that $\varphi:~A \simeq B$, with an arrow $f:A \to \separated T$, 
  $$T_f = \mathring T_B.$$
  For this, we use 
  \begin{align*}
    T_f &= \mathrm{Coeq}_1 \left( 
          \xymatrix{
          \sum_{a,b:A} f a = f b \dar[r]{4pt}^-{\pi_1}_-{\pi_2} & A
  }
                                                                  \right)
    \\
    \mathring T_B &= \mathrm{Coeq}_1 \left( 
                    \xymatrix{\sum_{a,b:B} \modal (a=b) \dar[r]{4pt}^-{\pi_1}_-{\pi_2}
                                                                & B
                                                                  \right)
  \end{align*}

%   Let $k:\nat$, we have to show that 
%   \begin{align*}
%     &\sum_{t:T^k} (\mu_T t_1 = \mu_T t_2 \land \cdots
%       \land \mu_T t_{k-1} \mu_T t_k)\\
%     &= \sum_{t:T^k} \modal_n (t_1 = t_2 \land \cdots
%       \land t_{k-1} = t_k)
%   \end{align*}
% By induction on $k$, and the preservation of products by $\modal_n$, it
% suffices to show that for any $a,b:T$, $\modal_n (a=b) = (\mu_T a =
% \mu_T b)$. By univalence, we want arrows in both ways, forming an
% equivalence.
% \begin{itemize}
% \item Suppose $p : (\mu_T a = \mu_T b)$. Then projecting $p$ along
%   first components yields $q : \prod_{t:T} \modal_n(a=t) = \modal_n (b=t)
%   $.
%   Taking for example $t=b$, we deduce $\modal_n (a=b) = \modal_n(b=b)$,
%   and the latter is inhabited by $\eta_{b=b} 1$.
% \item Suppose now $p : \modal_n(a=b)$. Let $\iota$ be the first
%   projection from $\separated T \to (T \to \Type_n^\modal)$. $\iota$ is
%   an embedding, thus it suffices to prove $\iota (\mu_T a) = \iota
%   (\mu_T b)$, \ie $\prod_{t:T}\modal_n (a=t) = \modal_n (b=t)$. The latter
%   remains true by univalence.
% \end{itemize}
% The fact that these two form an equivalence is technical, we refer to
% the formalization for an explicit proof.
\end{proof}
